---
title: "ST558 - Project 2"
author: "Tyler Pollard & Lucy Yin"
output: 
 html_document:
   toc: true
   toc_depth: 3
params: 
      weekday: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```
# Required Packages
```{r}
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(caret)
```

# Introduction (lucy)
  
  
# Data (lucy)
```{r data - read in}
# read in data
hour.data <- read_csv("data/hour.csv") %>% as_tibble()
day.data <- read_csv("data/day.csv") %>% as_tibble()
params$weekday
```

```{r data - change type}
# correct the variable types
hour.data$season <- factor(hour.data$season)
hour.data$yr <- factor(hour.data$yr)
hour.data$mnth <- factor(hour.data$mnth)
hour.data$hr <- factor(hour.data$hr)
hour.data$holiday <- factor(hour.data$holiday)
hour.data$weekday <- factor(hour.data$weekday)
hour.data$workingday <- factor(hour.data$workingday)
hour.data$weathersit <- factor(hour.data$weathersit)

day.data$season <- factor(day.data$season)
day.data$yr <- factor(day.data$yr)
day.data$mnth <- factor(day.data$mnth)
day.data$holiday <- factor(day.data$holiday)
day.data$weekday <- factor(day.data$weekday)
day.data$workingday <- factor(day.data$workingday)
day.data$weathersit <- factor(day.data$weathersit)
```

```{r data - create total data}
# add in a new variable before merging
hour.data <- mutate(hour.data, type = "hour")
day.data <- mutate(day.data, type = "day", hr = NA) %>% select(instant, dteday, season, yr, mnth, hr, everything())

# merge to create complete list of hour/day data
total.data <- rbind(hour.data, day.data)
```

```{r data - filter parameter}
# filter out to one specific day of the week
hour.data <- hour.data %>% filter(weekday == params$weekday)
day.data <- day.data %>% filter(weekday == params$weekday)
total.data <- total.data %>% filter(weekday == params$weekday)
```

```{r data - split training/test}
# splitting data into training & test sets
set.seed(7)
train <- sample(1:nrow(hour.data), size = nrow(hour.data)*0.7)
test <- dplyr::setdiff(1:nrow(hour.data), train)
training.data <- hour.data[train, ]
test.data <- hour.data[test, ]
```

# Summarization (both)
## Contingency Tables
### Weather Situation
```{r contigency table - weathersit}
#levels(day.data$weathersit) <- list("1" = "1", "2" = "2", "3" = "3", "4" = "4")
levels(day.data$weathersit) <- list(
  "Clear, Few clouds, Partly cloudy, Partly cloudy" = "1",
  "Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist" = "2",
  "Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds" = "3",
  "Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog" = "4")
kable(t(table(day.data$weathersit)))
```

## Summary Tables
### Adjusted Temperature
```{r summary table - atemp}

```
  
# Modeling
## Linear Regression Model
### What is Linear Regression Model (tyler)
  
### Linear Regression Model (tyler)

### Linear Regression Model (lucy)
  
## Ensemble Tree Model

### Random Forest Model (tyler)

### Boosted Tree Model (lucy)
  
# Comparison (either)
