---
title: "ST558 - Project 2"
author: "Tyler Pollard & Lucy Yin"
output: 
 html_document:
   toc: true
   toc_depth: 3
params: 
      weekday: monday
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
options(scipen = 999)
```
# Required Packages
```{r packages, echo=FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(caret)
library(gridExtra)
library(corrplot)
library(GGally)
```

# Introduction (lucy)
  
# Data (lucy)
```{r data - read in}
# read in data
hour.data <- read_csv("data/hour.csv") %>% as_tibble()
day.data <- read_csv("data/day.csv") %>% as_tibble()
```

```{r data - change type}
# correct the variaable types
hour.data$season <- factor(hour.data$season)
levels(hour.data$season) <- list(winter = 1, spring = 2, summer = 3, fall = 4)
hour.data$yr <- factor(hour.data$yr)
levels(hour.data$yr) <- list("2011" = 0, "2012" = 1)
hour.data$mnth <- factor(hour.data$mnth)
hour.data$hr <- factor(hour.data$hr)
hour.data$holiday <- factor(hour.data$holiday)
hour.data$weekday <- factor(hour.data$weekday)
levels(hour.data$weekday) <- list(monday = 1, tuesday = 2, wednesday = 3, thursday = 4, friday = 5, saturday = 6, sunday = 0)
hour.data$workingday <- factor(hour.data$workingday)
hour.data$weathersit <- factor(hour.data$weathersit)

day.data$season <- factor(day.data$season)
levels(day.data$season) <- list(winter = 1, spring = 2, summer = 3, fall = 4)
day.data$yr <- factor(day.data$yr)
levels(day.data$yr) <- list("2011" = 0, "2012" = 1)
day.data$mnth <- factor(day.data$mnth)
day.data$holiday <- factor(day.data$holiday)
day.data$weekday <- factor(day.data$weekday)
levels(day.data$weekday) <- list(monday = 1, tuesday = 2, wednesday = 3, thursday = 4, friday = 5, saturday = 6, sunday = 0)
day.data$workingday <- factor(day.data$workingday)
day.data$weathersit <- factor(day.data$weathersit)
```

```{r data - create total data}
# add in a new variable before merging
hour.data <- mutate(hour.data, type = "hour")
day.data <- mutate(day.data, type = "day", hr = NA) %>% select(instant, dteday, season, yr, mnth, hr, everything())

# merge to create complete list of hour/day data
total.data <- rbind(hour.data, day.data)
```

```{r data - filter parameter}
# filter out to one specific day of the week
hour.data <- hour.data %>% filter(weekday == params$weekday)
day.data <- day.data %>% filter(weekday == params$weekday)
total.data <- total.data %>% filter(weekday == params$weekday)
```

```{r data - split training/test}
# splitting data into training & test sets
set.seed(7)
train <- sample(1:nrow(day.data), size = nrow(day.data)*0.7)
test <- dplyr::setdiff(1:nrow(day.data), train)
day.training.data <- day.data[train, ]
day.test.data <- day.data[test, ]

hour.training.data <- hour.data[hour.data$dteday %in% day.training.data$dteday,]
hour.test.data <- hour.data[hour.data$dteday %in% day.test.data$dteday,]
```

# Summarization (both)
## Contingency Tables
### Year, Season and Count of Riders
```{r}

#table(day.training.data$season)
#table(day.training.data$yr)

kable(table(day.training.data$season, cut(day.training.data$cnt, breaks = 2)), caption = "Occurrences of # Range of Riders of a given Season")
kable(table(day.training.data$yr, cut(day.training.data$cnt, breaks = 2)), caption = "Occurrences of # Range of Riders of a given Year")
```
### Working Day and Count of Casual Riders
```{r}
levels(day.training.data$workingday) <- list("workday" = 1, "non-workday" = 0)
kable(table(day.training.data$workingday, cut(day.training.data$casual, breaks = 2)), caption = "Occurrences of # Range of Casual Riders of Workday vs. non-Workday")
```

## Summary Tables
### Humidity
```{r}
kable(t(c(summary(day.training.data$hum), St.Dev. = sd(day.training.data$hum))), digits = 2, caption = "Normalized Humidity")
kable(t(c(summary(day.training.data$hum*100), St.Dev. = sd(day.training.data$hum*100))), digits = 2, caption = "Raw Humidity")  
```
### Wind Speed
```{r}
kable(t(c(summary(training.data$windspeed), St.Dev. = sd(training.data$windspeed))), digits = 2, caption = "Normalized Wind Speed")
kable(t(c(summary(training.data$windspeed*67), St.Dev. = sd(training.data$windspeed*67))), digits = 2, caption = "Raw Wind Speed ")
```

## Density Plot
```{r}
ggplot(hour.training.data, aes(x = casual)) + 
  geom_density(alpha = 0.5, aes(fill = weathersit)) + 
  labs(title = "Density plot of casual riders",
       subtitle = "Specified by weather situation",
       x = "Count of Casual Riders",
       y = "Density") + 
  scale_fill_discrete(name = "Weather Situation", labels = c("Mostly Clear", "Mist", "Light Precip.", "Heavy Precip."))  
```

```{r}
ggplot(hour.training.data, aes(x = casual)) + 
  geom_density(alpha = 0.5, aes(fill = holiday)) +
  labs(title = "Density plot of casual riders",
       subtitle = "Specified by holiday or not",
       x = "Count of Casual Riders",
       y = "Density") +
  scale_fill_discrete(name = "Holiday?", labels = c("Non-Holiday", "Holiday"))    
```


## Scatter plots
```{r}
count.df <- hour.training.data
levels(count.df$mnth) <- list(January = 1, February = 2, March = 3, April = 4, May = 5, June = 6, July = 7, August = 8, September = 9, October = 10, November = 11, December = 12)
ggplot(count.df, aes(x = hr, y = cnt)) +
  geom_point(aes(col = workingday)) +
  facet_wrap(vars(mnth)) + 
  labs(title = "Count of riders for every hour of every month",
       subtitle = "Specified by workday or non-workday",
       x = "Hour of the Day",
       y = "Count of Riders") + 
  scale_color_discrete(name = "Working Day", labels = c("No", "Yes"))
```
## Box Plots
```{r}
ggplot(hour.training.data, aes(x = hr, y = cnt)) + 
  geom_boxplot() + 
  stat_summary(fun = mean, geom = "line", lwd = 0.8, aes(group = weathersit, col = weathersit)) + 
  labs(title = "Count of riders for every hr",
       subtitle = "Mean values based on weather situation",
       x = "Hour of the Day",
       y = "Count of Riders") + 
  scale_color_discrete(name = "Weather Situation", labels = c("Mostly Clear", "Mist", "Light Precip.", "Heavy Precip."))
```
## Correlation Plot 
```{r}
cor.variables <- hour.training.data %>% select(temp, atemp, hum, windspeed)
correlation <- cor(cor.variables, method = "spearman")
corrplot(correlation)
```
```{r}
subset.data <- data_frame(weathersit=day.training.data$weathersit, temp=day.training.data$temp, atemp=day.training.data$atemp,humidity=day.training.data$hum, windspeed=day.training.data$windspeed, casual=day.training.data$casual, registered=day.training.data$registered, total=day.training.data$cnt)
GGally::ggpairs(subset.data)
```
```{r}
subset.data2 <- data_frame(weathersit=hour.training.data$weathersit, temp=hour.training.data$temp, atemp=hour.training.data$atemp,humidity=hour.training.data$hum, windspeed=hour.training.data$windspeed, casual=hour.training.data$casual, registered=hour.training.data$registered, total=hour.training.data$cnt)
GGally::ggpairs(subset.data2)
```


# Modeling
## Linear Regression Model
### What is Linear Regression Model (tyler)
  
### Linear Regression Model (tyler)

### Linear Regression Model (lucy)
  
  
## Ensemble Tree Model
### Random Forest Model (tyler)

### Boosted Tree Model (lucy)
  
  
  
# Comparison (either)
